Floral preferences of mountain bumble bees are driven by trait matching but labile through elevation and time

### 2021-08-25 ms_2/ms_2_supp_methods.Rmd
preped ms for Ecol Lett

@@

>>>>

Patterns of preference through elevation and time were more complex than those of observed visitation and, by design, more often differed from patterns of floral abundance **(Figure 6)**. In some cases, though, patterns of preference mirrored patterns of floral availability, suggesting that bumble bees may exhibit functional responses [@Mysterud1998-rs] to floral abundance not entirely accounted for by null models that assume a linear proportionality between abundance and visitation.

@@

Broaching the topic of competition raises an issue that can confound any observational study of resource selection: how to disentangle the endogenous process of resource preference from the exogenous process of niche partitioning driven by competition [@Pianka1974-qk]. Empirical preference could reflect a latent process of resource preference, as we assume in our interpretation. Alternatively, though, visitation patterns and the preference inferred from them could be distorted by competitive forces. Intraspecific competition in a bumble bee community would tend to force individuals of a species to broaden their diet beyond what they would prefer under non-competitive conditions, while interspecific competition would be expected to cause a narrowing of diet breadth toward resources that can be exploited most efficiently. Thus, observed visitation patterns, even when calculated as abundance-relative preference, can only be interpreted in terms of *preference* to the extent that competition can be assumed to be negligible, and assumption must be justified on the basis of additional knowledge of the study system under consideration. Our study system in the Berchtesgadener Alps is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn]. We cannot, however, exclude the possibility that competitive forces distort, to some degree, the visitation patterns we observed. Even when competitive conditions do not obtain universally in a given system, competition may occur in a spatially and/or temporally localized way [@Ranta1981-vx], perhaps associated with extreme climatic events that cause reductions in floral resource avaulability [@Rasmont2012-zq; @Thomson2016-gq; @Phillips2018-me]. This calls to mind the insight of MacArthur1972-qz, that competition need not be frequent to have lasting ecological impacts, and it is likely that the "ghosts of competition past" [@Connell1980-dn] haunt the floral preferences exhibited by bumble bees today.
